Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-11-10 paper.tex
Small fixes in abstract and introduction

\begin{abstract}
  Splitting and merging data is a requirement for most parallel or
  distributed processing operations. Naive algorithms to split and
  merge 3D blocks from ultra-high resolution images perform very poorly,
  as a result of seek times. In contrast, naive algorithms to split and 
  merge 3D slabs perform optimally as seek time is significantly minimized.
  We introduce and analyze sequential algorithms (Clustered reads, Multiple reads, Clustered writes, and Multiple writes)
  that leverage memory buffering to
  address this issue. Clustered reads and Clustered writes, access
  image chunks only once, but they have to seek in the reconstructed
  image. Multiple reads and Multiple writes minimize seeks in the
  reconstructed image, but they access image chunks multiple
  times. Evaluation on a 3850x3025x3500 brain image shows that our
  algorithms perform similarly to the optimal configuration provided that enough memory is available. 
  Additionally, Multiple reads supports on-the-fly compression of the 
  merged image transparently but Clustered reads do not, due to their use of negative seeking. We
  conclude that splitting and merging large 3D images can be done
  efficiently without relying on complex data formats.
\end{abstract}

@@

% Problem definition
Three-dimensional images that exceed typical memory size are
increasingly found in a variety of disciplines. Big~Brain, for
instance, is a 3D histological image of the human brain that
represents 1~TB of raw data organized in 3600 planes at full
resolution and 76~GB at a 40-micrometer isotropic resolution commonly
used in neurosciences~\cite{amunts2013bigbrain}. Other examples found
in medical imaging, our primary domain of interest, include
high-resolution 3D electron microscopy (see, e.g.,
~\cite{bock2011network}) or micro- and
nano-tomography~\cite{10.1371/journal.pone.0035691}. As such images
would typically be processed on a computing cluster, possibly using
locality-aware file systems such as the Hadoop Distributed File System
(HDFS~\cite{shvachko2010hadoop}), software libraries are needed to
split and merge them efficiently, in particular to limit file
seeks. In this paper we introduce and compare a family of sequential
algorithms to split and merge images with reduced seeking.

We assume that the high-resolution image is split into chunks
representing 3D blocks or 3D slabs that fit in memory. A 3D block
consists of a stack of one or more 2D tiles (incomplete slices and/or columns) spanning sequentially occuring slices, whereas a 3D slab is 
a series of one of more contiguous 2D slices. A dataset such
as Big~Brain would perhaps be split into 125 chunks of 600~MB. The
decision to split an image into slabs or blocks, and the size of the
chunks, is up to the application. For instance, spatial filtering would
commonly require blocks, whereas slices would be preferable for acquisition artifact removal. 
Applications that process voxels individually, for instance
histogram computation or k-means clustering, could work on either
slices or blocks. Flexibility is thus required in the splitting
scheme.

>>>>

\begin{abstract}
  Splitting and merging data is a requirement for many parallel or
  distributed processing operations. Naive algorithms to split and
  merge 3D blocks from ultra-high resolution images perform very poorly,
  as a result of seek times. In contrast, naive algorithms to split and 
  merge 3D slabs perform optimally as seek time is significantly minimized.
  We introduce and analyze sequential algorithms (Clustered reads, Multiple reads, Clustered writes, and Multiple writes)
  that leverage memory buffering to
  address this issue. Clustered reads and Clustered writes, access
  image chunks only once, but they have to seek in the reconstructed
  image. Multiple reads and Multiple writes minimize seeks in the
  reconstructed image, but they access image chunks multiple
  times. Evaluation on a 3850x3025x3500 brain image shows that our
  algorithms perform similarly to the optimal configuration provided that enough memory is available. 
  Additionally, Multiple reads supports on-the-fly compression of the 
  merged image transparently but Clustered reads does not, due to its use of negative seeking. We
  conclude that splitting and merging large 3D images can be done
  efficiently without relying on complex data formats.
\end{abstract}

@@

% Problem definition
Three-dimensional images that exceed typical memory size are
increasingly found in a variety of disciplines. Big~Brain, for
instance, is a 3D histological image of the human brain that
represents 1~TB of raw data organized in 3600 planes at full
resolution and 76~GB at a 40-micrometer isotropic resolution commonly
used in neurosciences~\cite{amunts2013bigbrain}. Other examples found
in medical imaging, our primary domain of interest, include
high-resolution 3D electron microscopy (see, e.g.,
~\cite{bock2011network}) or micro- and
nano-tomography~\cite{10.1371/journal.pone.0035691}. As such images
would typically be processed on a computing cluster, possibly using
locality-aware file systems such as the Hadoop Distributed File System
(HDFS~\cite{shvachko2010hadoop}), software libraries are needed to
split and merge them efficiently, in particular to limit file
seeks. In this paper we introduce and compare sequential
algorithms to split and merge images with reduced seeking.

We assume that the high-resolution image is split into chunks
representing 3D blocks or 3D slabs that fit in memory. A 3D block
consists of a stack of one or more 2D tiles (incomplete slices or
incomplete columns) spanning contiguous slices, whereas a 3D slab is a
series of one of more complete contiguous 2D slices. A dataset such as
Big~Brain would perhaps be split into 125 chunks of 600~MB. The
decision to split an image into slabs or blocks, and the size of the
chunks, is up to the application. For instance, spatial filtering
would commonly require blocks, whereas slabs might be preferable for
acquisition artifact removal.  Applications that process voxels
individually, for instance histogram computation or k-means
clustering, could work on either slabs or blocks. Flexibility is thus
required in the splitting scheme.
